---
title: "01-RR-CATS"
author: "Anoff Nicholas Cobblah"
date: "July 31, 2018"
output: html_document
  html_document:
    number_sections: yes
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This script combines my Word Flagging and KWIC (tokenizer script) methods in order to create an interactive illustration of the frequency with which the terms "cat," "cats," "dog" or "dogs" are referenced in the novels of Charles Dickens. The goal is to answer two questions: which Dickens novel contains the most references to cats, and did Bob's death around 1862 have any influence on appearances of cats in Dickens's novels.

First, we set our parameters. There are several new things to take into account. Firstly, it is helpful to have three types of conlength values for Key Words in Context. Since the goal is to create a plot which allows KWIC to appear on rollover, 250 words is a bit much. So we have three conlengths of 250, 10, and 3 words.

**It's also important to note that since the creation of the KWIC script takes a long time, I've included a logical switch to make sure it will only create a new dataset if the "raw" data is changed. So the script is set not to run unless it senses a change in the number of bytes, or if the last created dataset is missing.**

```{r DecDCATSApp Parameters, eval=FALSE}
    DCATSlocation <- getwd()
    DCATSdoclocation <- paste0(getwd(),"/Novels")
    DCATSlongconlength <- 250
    DCATSshortconlength <- 3
    DCATSPOSconlength <- 10
    DCATScatsearchedtermlist <- c("cat")
    DCATSdogsearchedtermlist <- c("dog")
    DCATSsearchedtermlist <- c(DCATScatsearchedtermlist,DCATSdogsearchedtermlist)
    DCATSoutputlocation <- paste0(getwd(),"/WordFlagDataFrames")
    DCATSWordFlagdfPath <- paste0(DCATSoutputlocation,"/","DCATSWordFlagdf.txt")
    DCATSDocumentSize <- 32691727
```

To create the data frame compiling every reference to a term, or to upload the last version of that dataset, run the following script. Note that because this takes takes a long time to run, the dataset is always saved and called back up if possible.

```{r DecDCATSApp Word Flag, eval=FALSE}
      if(sum(file.info(list.files(DCATSdoclocation, all.files = TRUE, recursive = TRUE, full.names=TRUE))$size) == DCATSDocumentSize) {
        DCATSDataChange1 <- FALSE
        print("The data in the 'Documents' folder appears not to have changed.")
      }else{
        DCATSDataChange1 <- TRUE
        print("The data in the 'Documents' folder appears to have been changed. A new DCATSWordFlagdf will therefore be created. TO UPDATE THIS SCRIPT, PLEASE CHANGE THE DCATSDocumentSize TO REFLECT THE NEW SIZE OF THE DOCUMENTS.")
        }
      
      if(file.exists(DCATSWordFlagdfPath) == TRUE) {
        DCATSDataChange2 <- FALSE
        print("The previous DCATSWordFlagdf still exists.")
      }else{
        DCATSDataChange2 <- TRUE
        print("The previous DCATSwordFlagdf seems to have been moved or deleted.  A new DCATSWordFlag will therefore be created.")
        }

  if(DCATSDataChange1|DCATSDataChange2 == TRUE) {
  
      files <- list.files(path = DCATSdoclocation, pattern = "xml", full.names = TRUE) #creates vector of xml file names.
      if(file.exists(DCATSoutputlocation) == FALSE){dir.create(DCATSoutputlocation)}
      DCATSstemsearchedtermlist <- unique(wordStem(DCATSsearchedtermlist)) #lemmatizes the list of terms you want to search for.
      DCATSWordFlagmat <- matrix(,ncol=12,nrow=1)
      for (i in 1:length(files)) {
        fileName <- xml_text(read_xml(files[i]))
        ltoken <- tokenize_words(fileName, lowercase = TRUE, stopwords = NULL, simplify = FALSE)
        ltoken <- unlist(ltoken)
        stemltoken <- wordStem(ltoken) #this uses the Snowball library to lemmatize the entire text.
        textID <- i
        for (p in 1:length(DCATSstemsearchedtermlist)) {
          DCATSstemsearchedterm <- DCATSstemsearchedtermlist[p]
          for (j in 1:length(stemltoken)) {
              if (DCATSstemsearchedterm == stemltoken[j]) {
                if (j <= DCATSlongconlength) {longtempvec <- ltoken[(1:(j+DCATSlongconlength))]}
                if (j > DCATSlongconlength) {longtempvec <- ltoken[(j-DCATSlongconlength):(j+DCATSlongconlength)]}
                if (j <= DCATSshortconlength) {shorttempvec <- ltoken[(1:(j+DCATSshortconlength))]}
                if (j > DCATSshortconlength) {shorttempvec <- ltoken[(j-DCATSshortconlength):(j+DCATSshortconlength)]}
                if (j <= DCATSPOSconlength) {POStempvec <- ltoken[(1:(j+DCATSPOSconlength))]}
                if (j > DCATSPOSconlength) {POStempvec <- ltoken[(j-DCATSPOSconlength):(j+DCATSPOSconlength)]}
                TempTextName <- gsub(paste0(DCATSdoclocation,"/"),"",files[i]) #This grabs just the end of the file path.
                TempTextName <- gsub(".xml","",TempTextName) #This removes the .xml from the end of the name.
                temprow <- matrix(,ncol=12,nrow=1)
                colnames(temprow) <- c("Text", "Text_ID", "DCATSstemsearchedterm","Lemma","Lemma_Perc","KWIC","Total_Lemma","Date","Category","Short_KWIC","POS_KWIC","Current_Date")
                temprow[1,1] <- strsplit(TempTextName,"_")[[1]][3]
                temprow[1,2] <- textID
                temprow[1,3] <- DCATSstemsearchedterm
                temprow[1,4] <- j
                temprow[1,5] <- (j/length(stemltoken))*100
                temprow[1,6] <- as.character(paste(longtempvec,sep= " ",collapse=" "))
                temprow[1,7] <- length(stemltoken)
                temprow[1,8] <- strsplit(TempTextName,"_")[[1]][2]
                #Determining Category
                  if(DCATSstemsearchedterm %in% wordStem(DCATScatsearchedtermlist)) {temprow[1,9] <- "Cat-References"}
                  if(DCATSstemsearchedterm %in% wordStem(DCATSdogsearchedtermlist)) {temprow[1,9] <- "Dog-References"}
                temprow[1,10] <- as.character(paste(shorttempvec,sep= " ",collapse=" "))
                temprow[1,11] <- as.character(paste(POStempvec,sep= " ",collapse=" "))
                temprow[1,12] <- format(Sys.time(), "%Y-%m-%d")
                DCATSWordFlagmat <- rbind(DCATSWordFlagmat,temprow)
              }
          }
        }
        print(files[i]) #let's user watch as code runs for long searches
      }
      DCATSWordFlagmat <- DCATSWordFlagmat[-1,]
      DCATSWordFlagdf <- as.data.frame(DCATSWordFlagmat)
      write.table(DCATSWordFlagdf, DCATSWordFlagdfPath)
      DCATSWordFlagdf[1:5,]
  }else{
    print("Loading the previous dataset as DCATSWordFlagdf")
    DCATSWordFlagdf <- read.table(DCATSWordFlagdfPath)
  }
DCATSWordFlagdf$KWIC = as.character(DCATSWordFlagdf$KWIC)
      DCATSWordFlagdf$Text = as.character(DCATSWordFlagdf$Text)
DCATSWordFlagdf
```

We should note that there are going to be some false references here. For instance, two of the references to cats in Bleak House will be to Monsieur Le Cat.

We can then add up the values in DCATSWordFlagdf to make a table of the frequency of cat and dog references, DCATSFreqmat. This step is important because we also introduce a new column: "Sample_KWIC", which randomly chooses one example of Key Words in Context.
```{r,  eval=FALSE}
  # Adding values from DCATSWordFlagdf together to get a matrix of normalized frequencies for each category, as DCATSFreqmat
      DCATSWordFlagcatdf <- DCATSWordFlagdf[grep("Cat-References",DCATSWordFlagdf$Category),]
      DCATSWordFlagdogdf <- DCATSWordFlagdf[grep("Dog-References",DCATSWordFlagdf$Category),]
      DCATSFreqmat <- matrix(,ncol=8,nrow=1)
      files <- list.files(path = DCATSdoclocation, pattern = "xml", full.names = TRUE) #creates vector of xml file names.
      for (i in 1:length(files)) {
        TempFileName <- gsub(paste0(DCATSdoclocation,"/"),"",files[i]) #This grabs just the end of the file path.
        TempFileName <- gsub(".xml","",TempFileName) #This removes the .xml from the end of the name.
        TempTextName <- strsplit(TempFileName,"_")[[1]][3] #This just grabs the title
        tempcatdf <- DCATSWordFlagcatdf[grep(TempTextName,DCATSWordFlagcatdf$Text),]
        tempdogdf <- DCATSWordFlagdogdf[grep(TempTextName,DCATSWordFlagdogdf$Text),]
        TempDate <- strsplit(TempFileName,"_")[[1]][2]
        TempLength <- tempcatdf$Total_Lemma[1]
        temprows <- matrix(,ncol=8,nrow=2)
        colnames(temprows) <- c("Text", "Text_ID","Date","Category","Frequency","Total_Lemma","Normalized_Freq","Sample_KWIC")
        temprows[1:2,1] <- as.character(TempTextName)
        temprows[1:2,2] <- i
        temprows[1:2,3] <- as.character(TempDate)
        temprows[1,4] <- "Cat-References"
        temprows[2,4] <- "Dog-References"
        temprows[1,5] <- nrow(tempcatdf)
        temprows[2,5] <- nrow(tempdogdf)
        temprows[1:2,6]<- as.character(TempLength)
        temprows[1,7] <- (as.numeric(temprows[1,5])/as.numeric(temprows[1,6]))*100
        temprows[2,7] <- (as.numeric(temprows[2,5])/as.numeric(temprows[2,6]))*100
        #temprows[1,8]
          if(nrow(tempcatdf) > 0){temprows[1,8] <- as.character(sample(tempcatdf$Short_KWIC,1))}else{temprows[1,8] <- NA}
        #temprows[2,8]
          if(nrow(tempdogdf) >0) {temprows[2,8] <- as.character(sample(tempdogdf$Short_KWIC,1))}else{temprows[2,8] <- NA}
        DCATSFreqmat <- rbind(DCATSFreqmat,temprows)
      }
      DCATSFreqmat <- DCATSFreqmat[-1,]
      DCATSFreqdf <- as.data.frame(DCATSFreqmat)
      
      DCATSFreqdf

```

We can then answer questions using DCATSFreqdf.  For instance, we can see that Dickens references dogs far more often than cats in his novels.

```{r DCATSFreqmat by date,  eval=FALSE}
      
  # Visualizing DCATSFreqdf
      p <- ggplot(DCATSFreqdf, aes(y = as.numeric(as.character(Normalized_Freq)), x = as.numeric(as.character(Date)), color = Category))
      pg <- geom_point(size=1,pch = 16)
      pl <- p + pg + labs(x = "Date", y = "Normalized Frequency (% of Words in Text)", title = "Appearances of 'cat' and 'dog' within Dickens's Novels")
      pl
```

Or we can create an interactive graph that shows us an example key word in context from that year on rollover.

```{r DCATSFreqmat by date interactive,  eval=FALSE}
      
  # Visualizing DCATSFreqdf
      p <- ggplot(DCATSFreqdf, aes(y = as.numeric(as.character(Normalized_Freq)), x = as.numeric(as.character(Date)), color = Category, label = Sample_KWIC))
      pg <- geom_point(size=1,pch = 16)
      pl <- p + pg + labs(x = "Date", y = "Normalized Frequency (% of Words in Text)", title = "Appearances of 'cat' and 'dog' in Dickens's Novels")
      ggplotly(pl)

```


Or we can create an interactive graph that shows us the novel we're looking at on rollover. Bleak House and Edwin Drood are apparently the only novels where references to cats approach references to dogs.

```{r DCATSFreqmat by date interactive text,  eval=FALSE}
      
  # Visualizing DCATSFreqdf
      p <- ggplot(DCATSFreqdf, aes(y = as.numeric(as.character(Normalized_Freq)), x = as.numeric(as.character(Date)), color = Category, label = Text))
      pg <- geom_point(size=1,pch = 16)
      pl <- p + pg + labs(x = "Date", y = "Normalized Frequency (% of Words in Text)", title = "Appearances of 'cat' and 'dog' in Dickens's Novels")
      ggplotly(pl)

```

We can verify this by looking at the raw frequencies. Cats and dogs are most equal in Edwin Drood. By far the biggest discrepancy between cats and dogs is in Oliver Twist.

```{r DCATSFreqmat by novel,  eval=FALSE}
      
  # Visualizing DCATSFreqdf
      p <- ggplot(DCATSFreqdf, aes(y = as.numeric(as.character(Frequency)), x = as.numeric(as.character(Date)), color = Category, label = Text))
      pg <- geom_point(size=1,pch = 16)
      pl <- p + pg + labs(x = "Date", y = "Frequency (# of References in Text)", title = "Appearances of 'cat' and 'dog' in Dickens's Novels")
      ggplotly(pl)

```

We can make this EVEN CLEARER by instead visualizing the difference between cats and dogs in each novel. So cat lovers should either turn to Bleak House or Edwin Drood
```{r}
difference <- matrix(,ncol=3,nrow=1)        
files <- list.files(path = DCATSdoclocation, pattern = "xml", full.names = TRUE) #creates vector of xml file names.
      for (i in 1:length(files)) {
        TempFileName <- gsub(paste0(DCATSdoclocation,"/"),"",files[i]) #This grabs just the end of the file path.
        TempFileName <- gsub(".xml","",TempFileName) #This removes the .xml from the end of the name.
        TempTextName <- strsplit(TempFileName,"_")[[1]][3] #This just grabs the title
        TempDate <- strsplit(TempFileName,"_")[[1]][2]
        tempdf <- DCATSFreqdf[grep(TempTextName,DCATSFreqdf$Text),]
        temprows <- matrix(,ncol=3,nrow=1)
        colnames(temprows) <- c("Text","Date","DogCatDif")
        temprows[1,1] <- as.character(TempTextName)
        temprows[1,2] <- as.character(TempDate)
        temprows[1,3] <- as.numeric(as.character(tempdf[grep("Dog-References",tempdf$Category),][1,5]))-as.numeric(as.character(tempdf[grep("Cat-References",tempdf$Category),][1,5]))

        difference <- rbind(difference,temprows)
      }
      difference <- difference[-1,]
      differencedf <- as.data.frame(difference)
      differencedf
      
      p <- ggplot(differencedf, aes(y = as.numeric(as.character(DogCatDif)), x = as.factor(Text)))
      pg <- geom_bar(stat="identity")
      pl <- p + pg + labs(y = "Dog References - Cat References", x = "Text", title = "Comparison of 'cat' and 'dog' in Dickens's Novels") + scale_x_discrete(limits=differencedf$Text) + coord_flip()
      pl
```

We can also visualize the terms which most frequently occur around the search terms in the two categories within this corpus. Interestingly, cats seem to be associated with women and dogs with men.
```{r DCATS dog/cat associations,  eval=FALSE}
corpus <- corpus(DCATSWordFlagdf, 
                 docid_field="Text", 
                 text_field="KWIC")
group_DCATSWordFlagdfm <- dfm(corpus, remove=c(stopwords("en"),"cat","cats","dogs","dog"), remove_punct=TRUE, remove_numbers = TRUE, groups="Category")
textplot_wordcloud(group_DCATSWordFlagdfm,max.words=50, colors = RColorBrewer::brewer.pal(8,"Dark2"), comparison=TRUE)


```

Finally, we can run a very rudimentary qualitative sentiment analysis by looking at JUST the adjectives which appear around the term (for instance, within a 10 word range on either side).  This also requires the use of the coreNLP library, which can take a long time to install and initialize.

**IMPORTANT NOTE: Since creating a Word Flag matrix can take a nontrivial amount of time for larger corpuses, this script is designed only to run the program to create a new DCATSWordFlagdf if there is a change to the dataset in folder "Documents" or if the previous DCATSKWICPOScatdf and DCATSKWICPOSdogdf has been deleted.**

First, we add a few more parameters determining what the output will be saved as.
```{r DECDCATSPOSApp parameter,  eval=FALSE}
    DCATSKWICPOScatdfPath <- paste0(DCATSoutputlocation,"/","DCATSKWICPOScatdf.txt")
    DCATSKWICPOSdogdfPath <- paste0(DCATSoutputlocation,"/","DCATSKWICPOSdogdf.txt")
```

Then we run a script to create dataset which identify and mark each of the adjectives and adverbs in our Key Words in context (the 10 word range one, as otherwise this quickly gets out of hand.)

```{r DECDCATSPOSApp,  eval=FALSE}
   if(file.exists(DCATSKWICPOScatdfPath)&file.exists(DCATSKWICPOSdogdfPath) == TRUE) {
        DCATSDataChange3 <- FALSE
        print("The previous DCATSWordFlagdf still exists.")
      }else{
        DCATSDataChange3 <- TRUE
        print("The previous DCATSKWICPOScatdf or DCATSKWICPOSdogdf seems to have been moved or deleted.  A new DCATSKWICPOSdf will therefore be created.")
        }
  
  if(DCATSDataChange1|DCATSDataChange3 == TRUE) {
    #we run part of speech tagging on each of these KWIC and draw out just the adjectives, and sum up the numbers.
      #We do this for the cat References data.
        ADJADVcatdf <- data.frame( Var1=character(),Freq=numeric())
        for(i in 1:nrow(DCATSWordFlagcatdf)) {
          tempstring <- as.character(DCATSWordFlagcatdf$POS_KWIC[i])
          anno <- annotateString(tempstring)
          token <- getToken(anno)
          ut <- universalTagset(token$POS)
          index <- c(which(ut=="ADJ"), which(ut=="ADV"))
          temptable <- table(token$lemma[index])
          ADJADVcatdf <- rbind(ADJADVcatdf,as.data.frame(temptable))
          print(paste0(i," out of ",nrow(DCATSWordFlagcatdf)))
        }
        ADJADVcatdf <- aggregate(ADJADVcatdf$Freq, b=list(Category=ADJADVcatdf$Var1), FUN=sum)
        DCATSKWICPOScatdf <- ADJADVcatdf[order(ADJADVcatdf$x, decreasing=TRUE),]  #reordering the matrix
        write.table(DCATSKWICPOScatdf, DCATSKWICPOScatdfPath)
        
      #And for the dog References data.
        ADJADVdogdf <- data.frame( Var1=character(),Freq=numeric())
        for(i in 1:nrow(DCATSWordFlagdogdf)) {
          tempstring <- as.character(DCATSWordFlagdogdf$POS_KWIC[i])
          anno <- annotateString(tempstring)
          token <- getToken(anno)
          ut <- universalTagset(token$POS)
          index <- c(which(ut=="ADJ"), which(ut=="ADV"))
          temptable <- table(token$lemma[index])
          ADJADVdogdf <- rbind(ADJADVdogdf,as.data.frame(temptable))
          print(paste0(i," out of ",nrow(DCATSWordFlagdogdf)))
        }
        ADJADVdogdf <- aggregate(ADJADVdogdf$Freq, b=list(Category=ADJADVdogdf$Var1), FUN=sum)
        DCATSKWICPOSdogdf <- ADJADVdogdf[order(ADJADVdogdf$x, decreasing=TRUE),]  #reordering the matrix
        write.table(DCATSKWICPOSdogdf, DCATSKWICPOSdogdfPath)
  }else{
    print("Loading the previous datasets as DCATSKWICPOScatdf and DCATSKWICPOSdogdf")
    DCATSKWICPOScatdf <- read.table(DCATSKWICPOScatdfPath)
    DCATSKWICPOSdogdf <- read.table(DCATSKWICPOSdogdfPath)
  }
    DCATSKWICPOScatdf
    DCATSKWICPOSdogdf
```

Finally, we visualize the top 25 adjectives and adverbs in these KWIC sets. In this case, we don't learn very much, unfortunately.

```{r DECDCATSPOSApp Visualized,  eval=FALSE}
        TopADJADVcatdf <- DCATSKWICPOScatdf[1:25,]
        TopADJADVcatdf$Category <- factor(TopADJADVcatdf$Category, levels = TopADJADVcatdf$Category[order(TopADJADVcatdf$x)])
        TopADJADVdogdf <- DCATSKWICPOSdogdf[1:25,]
        TopADJADVdogdf$Category <- factor(TopADJADVdogdf$Category, levels = TopADJADVdogdf$Category[order(TopADJADVdogdf$x)])
    
        #Then we visualize the top 25 adjectives and adverbs for dog and cat References.
           p1 <- ggplot(TopADJADVcatdf, aes(y = as.numeric(as.character(x)), x = (Category)))
           p2 <- geom_bar(stat="identity") 
           p3 <- p1 + p2 + labs(x = "Adjective/Adverb near cat References", y = "Frequency", title = "Common Adjectives and Adverbs near cat References \nwithin Dickens's Novels")
           pl1 <- p3+coord_flip()
          
            p4 <- ggplot(TopADJADVdogdf, aes(y = as.numeric(as.character(x)), x = (Category)))
           p5 <- geom_bar(stat="identity") 
           p6 <- p4 + p5 + labs(x = "Adjective/Adverb near Cat References", y = "Frequency", title = "Common Adjectives and Adverbs near Dog References \nwithin Dickens's Novels")
           pl2 <- p6+coord_flip()
           {print(pl1)
           print(pl2)}
```
